### 2020-04-10 alpibot.tex
Subo un monton de info.  Parte de la intro, subo todas las imagenes de los resultados, agrego las tablas, agrego el clinical assessment y parte de la discusión.  Falta pero es un avance.

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{algorithm2e} % For algorithms
\usepackage{amsmath, amssymb, amsthm} % Mathematical packages
\hypersetup{hidelinks=true}    
\usepackage{algorithmic}

%\usepackage{fancyhdr}
%\usepackage{array}
%\usepackage[rightcaption]{sidecap}
%\usepackage{wrapfig}
%\usepackage{float}
%\usepackage[labelfont=bf]{caption} % bold text for captions
%\usepackage[para]{threeparttable} % fancy tables, check these before you use them
%\usepackage{url}
%\usepackage[table,xcdraw]{xcolor}
\usepackage{tikz}
%\usepackage{pgfplots}
%\pgfplotsset{width=10cm,compat=1.9}
%\usepgfplotslibrary{external}
%\tikzexternalize
%\usepackage{svg}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
%\usepackage[pdftex]{graphicx}
\usepackage{subfig}

  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{booktabs}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{A Control Strategy for a Tethered Follower Robot for Pulmonary Rehabilitation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Bianchi Luciano, Buzniak Esteban, Ramele~Rodrigo,~\IEEEmembership{Member,~IEEE,}
        and Santos~Juan~Miguel% <-this % stops a space
\thanks{R. Ramele  and J.M.Santos are with the Department
of  Computer Engineering, Instituto Tecn{o'}gico de Buenos Aires (ITBA), Ciudad de Buenos Aires,  Argentina e-mail: rramele@itba.edu.ar}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Biomedical Engineering}%
{Bianchi \MakeLowercase{\textit{et al.}} Controller Design for a Tethered Follower Robot for Pulmonary Rehabilitation}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Patients that suffer Chronic Obstructive Pulmonary Disease (COPD) underwent a procedure called Pulmonary Rehabilitation that helps them to improve disease prognosis.  During Pulmonary Rehabilitation procedures patients require external oxygen assistance.  The oxygen tank cannot be carried by the patient and some external assistance is required.   In this work a basic robot follower is proposed to carry the oxygen tank based on a differential tethered scheme.  Two algorithms are proposed in a very basic configuration.  
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
robotics, tethered, COPD
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.

%The EEG is traditionally analyzed in terms of temporal waveforms at certain channels, looking at power of rhythms in the spontaneous EEG, at amplitude and latency of the peaks and troughs in event- related potentials (ERPs), or at particular grapho-elements in patho- logical or sleep stages.

\IEEEPARstart{C}{hronic} Obstructive Pulmonary Disease (COPD) is an umbrella term that describe several pulmonary affections characterized by a eskeletomuscular atrophy~\cite{Kocsis2016,Wu2012}.  In order to carve these after effects a Pulmonary Rehabilitation procedure is a viable treatment for these patients.  During these procedures patients underwent controlled walking activities and physical exercises under the supervision of a physical therapist.  These patients are characterized by a severe low saturation illness and they require effective oxygen supply, particularly when performing physical activity~\cite{Celli2014}.  Hence, patients require to carry with them an oxygen tank for the oxygenotherapy assistance.  However, their own condition prevent them with the ability to precisely carry the often bulky external tank.  This situation entails a pragmatic solution to provide an additional physical therapist solely for the purpose of carrying the oxygen tank. Alternatively, they require to devise a customized wheelchair to allow them to carry the oxygen tank on top of it, leading to a cumbersome situation for the physical therapist because s/he had to push and maneuver the wheelchair, and at the same time, take care of the patient and perform the rehabilitation treatment.

The scenario may be aggravated by the fact that this procedure is performed on a rehabilitation gym that could be potentially crowded with several patients, physiologist and physical therapists.

An alternative solution is to use an assitive ground service robot~\cite{Neto2015} to carry the oxygen tank, while following the patient in a leader-follower configuration.  There are two reasons that support the initial viability of this idea.  First,  the rehabilitation gym is a constrained environment where this problem can be tackled by an Unmanned Ground Vehicle (UGV).  On the other hand, the range of movements performed by the patient during the Rehabilitation Procedure is highly predictable by the treatment.  At the same time,  the global robotic research community persuade the development of robotic affordable solutions to the social and health-related worldwide problems~\cite{Khamis2019}.

For the implementation of the leader-follower strategy, several solutions have been proposed, including SLAM solutions, vision-based systems or based on electromagnetical beacons~\cite{Islam2019}.   The work presented by \cite{Ortlieb2016} explores a differential tethered robotic system to perform camera-based gait analysis of the leader.  For COPD Pulmonary Rehabilitaiton procedures, the patient is already umbilicaly linked to the oxygen tank via the breathing cannula.  Hence, a robotic solution can exploit this circumstance to perform the \textit{following} mechanism based on a tethered controller. Tethered robots have been extensively researched in robotics~\cite{Ahn2006}.  They offer a very simple solution to some common navigation problems, and they can be very effective in robot-to robot interaction, collaborative robotics, or while interacting with humans in Human Robot Interfaces~\cite{Rekleitis2001,Hirata2009,Ferrin2010}.

On the other hand, several assistance devices for COPD treatments have been proposed.  Particularly relevant are novel telemedice applications to enhance complementary rehabiliation exersice at home that can track biological markers for those patients~\cite{Wu2012}.  The work presented here follows the line pointed by \cite{Endo2015}.  Authors studied the use of a single thread tethered follower robot for home oxygen therapy, and compared two different control algorithms and their effectiveness to mimic the leader path and to avoid obstacles.  However their approach, focuses on the usage of the device exclusively for home therapy, and not within the context of a Pulmonary Rehabilitation procedure performed by medical personnel on medical installations.

Hence, this work aims to provide an initial study on the controlling mechanism of a differential tethered robot that implements the leader-follower configuration on a Pulmonary Rehabilitation procedure.  This document unrolls as follows.  Section~\ref{materials} describes poses the problem and the solution design, Section~\ref{experimental} document the experimental protocol to perform the solution assessment on a simulation and on a real world scenario.  Results and Discussions are described in~\ref{results} and \ref{discussion}. The Clinical Assessment performed jointly with medical personnel is tackled in Section~\ref{clinical}.  Finally, conclusions are exposed in the remaining section~\ref{conclusion}.

\section{Materials and Methods}
\label{materials}

In order to prioritize several iterations that could be verified by key stakeholders, we developed a simple robotic prototype constructed on Internet of Robotic Things~\cite{Simoens2018}.  This discipline is part of what is being called 4th industrial revolution and is reshaping many aspects of global manufacturing, including robotic and healthcare.

\subsection{Solution Design}

The medical community is frequently skeptical to technological solutions for medical personel~\cite{Gaggioli2006,Fasola2012,Cherubini2008,Wolpaw2012BraincomputerIP}.  At the same time, development process as design thinking and other similar tendencies prioritize rapid development, prototyping that can bring quickly feedback from real users about the drawbacks and improving opportunities.  In this context, it is important to perform rapid prototyping.

The proposed solution is a tethered robot follower, using a pair of strings.  The following mechanism is implemented by means of measuring encoders on two different reels located according to figure~\ref{fig:components}.

The driving mechanism proposed is a differential wheel robot.

\subsection{Hardware}

In order to achieve quick prototyping cycles, we used extensively tools and modules to the IoRT wave.  Frames are constructed from aluminum extrusions produced by Makeblock (Shenzhen, China).

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/alpibot1.jpg}
\caption{This robot prototype.}
\label{fig:performance}
\end{figure}

Motors are in-wheel configuration and they provide optical encoding.  A microcontroller is used to implement motor controlling and encoder processing.  This is based on the popular Arduino platform.  On top of it the Adafruit Shield v2.  This Arduino bridge can be used to control simultaneously four DC motors.

The microcontroller is connected through serial connection to the main SBC RaspberryPi Model 3B+.  The main controlling algorithm run on this board, it connects to wifi and provides telemetry and the abiilty to receive remote commands by means a very simple UDP command interface.

Reels were designed from PVC extrusions based on figure.  They are glued to regular DC motors scavenged from old compact discs.  Hardware specifications are described on table \ref{tab:specifications}.

\begin{figure}[]
    \centering
    \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
    \begin{tikzpicture}[x=0.50pt,y=0.50pt,yscale=-1,xscale=1]
    %uncomment if require: \path (0,1774); %set diagram left start at 0, and has height of 1774
    %Shape: Rectangle [id:dp3163749047300972] 
    \draw   (341,1289.29) -- (356.46,1289.29) -- (356.46,1484.1) -- (341,1484.1) -- cycle ;
    %Straight Lines [id:da006917544597343617] 
    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][line width=0.75]    (309.46,1464.29) -- (126.19,1383.9) ;

    %Straight Lines [id:da6556986586628986] 
    \draw [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 230; green, 164; blue, 164 }  ,fill opacity=1 ][line width=0.75]    (310.46,1310.29) -- (126.19,1383.9) ;

    %Shape: Rectangle [id:dp7070304476532343] 
    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (309.5,1289) -- (388,1289) -- (388,1324.67) -- (309.5,1324.67) -- cycle ;
    %Shape: Ellipse [id:dp4674581553313458] 
    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (110.01,1383.9) .. controls (110.01,1374.53) and (117.25,1366.94) .. (126.19,1366.94) .. controls (135.12,1366.94) and (142.37,1374.53) .. (142.37,1383.9) .. controls (142.37,1393.28) and (135.12,1400.87) .. (126.19,1400.87) .. controls (117.25,1400.87) and (110.01,1393.28) .. (110.01,1383.9) -- cycle ;
    %Shape: Rectangle [id:dp46270765707902994] 
    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (308.5,1448.5) -- (387,1448.5) -- (387,1484.17) -- (308.5,1484.17) -- cycle ;
    %Shape: Circle [id:dp1556138070188774] 
    \draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (497.83,1383.83) .. controls (497.83,1379.51) and (501.34,1376) .. (505.67,1376) .. controls (509.99,1376) and (513.5,1379.51) .. (513.5,1383.83) .. controls (513.5,1388.16) and (509.99,1391.67) .. (505.67,1391.67) .. controls (501.34,1391.67) and (497.83,1388.16) .. (497.83,1383.83) -- cycle ;
    %Straight Lines [id:da9759075394788362] 
    \draw    (362.11,1349.34) -- (505.67,1383.83) ;

    %Straight Lines [id:da0292648785573405] 
    \draw    (362.11,1423.58) -- (429.74,1404.86) -- (505.67,1383.83) ;

    %Rounded Rect [id:dp19413777874284577] 
    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 184; green, 233; blue, 134 }  ,fill opacity=1 ] (336,1423.58) .. controls (336,1422.16) and (337.16,1421) .. (338.58,1421) -- (359.53,1421) .. controls (360.96,1421) and (362.11,1422.16) .. (362.11,1423.58) -- (362.11,1431.34) .. controls (362.11,1432.76) and (360.96,1433.92) .. (359.53,1433.92) -- (338.58,1433.92) .. controls (337.16,1433.92) and (336,1432.76) .. (336,1431.34) -- cycle ;
    %Rounded Rect [id:dp9015992902460356] 
    \draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 184; green, 233; blue, 134 }  ,fill opacity=1 ] (336,1341.58) .. controls (336,1340.16) and (337.16,1339) .. (338.58,1339) -- (359.53,1339) .. controls (360.96,1339) and (362.11,1340.16) .. (362.11,1341.58) -- (362.11,1349.34) .. controls (362.11,1350.76) and (360.96,1351.92) .. (359.53,1351.92) -- (338.58,1351.92) .. controls (337.16,1351.92) and (336,1350.76) .. (336,1349.34) -- cycle ;

    % Text Node
    \draw  [draw opacity=0][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ]  (478.5,1350) -- (535.5,1350) -- (535.5,1374) -- (478.5,1374) -- cycle  ;
    \draw (507,1362) node    {$subject$};
    % Text Node
    \draw (302,1344) node    {$left\ reel$};
    % Text Node
    \draw (299,1425) node    {$right\ reel$};
    % Text Node
    \draw (125,1414) node    {$castor\ wheel$};
    % Text Node
    \draw (422,1348) node  [rotate=-12.55]  {$left\ thread$};
    % Text Node
    \draw (423,1422) node  [rotate=-344.79]  {$right\ thread$};
    % Text Node
    \draw (348,1275) node    {$left\ wheel$};
    % Text Node
    \draw (348,1496) node    {$right\ wheel$};
    \end{tikzpicture}
    \caption{Components of the robotic vehicle and following mechanism}
    \label{fig:components}
\end{figure}{}

\begin{figure*}[!t]
\centering
\subfloat[Case I]{\includegraphics[width=15cm]{images/alpibot2.jpg}\label{subject8}}
\caption{Robot prototype built to evaluate the real world scenario of described following mechanisms for the tethered robot.}
\label{fig:p300templates}
\end{figure*}

\subsection{Active Reel Spring}

As previously mentioned, an active spring mechanism is also put in place to keep the threads tense at all times and measure the distance to the subject precisely. This active spring mechanism can be implemented by just activating both reel motors at all times to keep the threads tense. However, in order to extend the useful life of the reel motors, and to save battery, an algorithm to intelligently activate and deactivate the motors was developed.

The algorithm works as follows:

\begin{enumerate}
    \item While wheels are moving, retract reels.
    \item If wheels stop moving, wait for \textit{reel wait time} seconds, then retract reels. 
    \item Retract reels until the reel encoders values have not changed during \textit{reel retract time} seconds.
    \item If wheels started moving or the encoder values have changed while retracting, start the \textit{reel retract time} countdown again.
\end{enumerate}{}

%Here is the pseudocode that implements that logic. This function is executed in every run loop in the Arduino module:

%\begin{algorithm}[]
%\label{alg:active_spring}
%\caption{Active spring algorithm}
%\begin{algorithmic}
%
%\STATE $currentReelPulses \gets readPulses()$
%
%\IF{$retracting$}
%    \IF{$lastReelPulses != currentReelPulses$}
%    \STATE $lastReelStart \gets currentTime()$
%    \ENDIF
%    \IF{\NOT $ wheelsMoving $ \AND $ currentTime() - lastReelStart > rrt$}
%    \STATE $retracting \gets \FALSE$
%    \STATE $lastReelEnd \gets currentTime()$
%    \ENDIF
%\ELSE
%    \IF{$wheelsMoving$}
%        \STATE $retracting \gets \TRUE$
%        \STATE $lastReelStart \gets currentTime()$
%    \ENDIF
%    \IF{$ currentTime() - lastReelEnd > rwt $}
%        \STATE $retracting \gets \TRUE$
%        \STATE $lastReelStart \gets currentTime()$
%    \ENDIF
%\ENDIF
%
%\STATE $lastReelPulses \gets currentReelPulses$
%
%\IF{$retracting$}
%    \STATE \textit{Activate reel motors}
%\ELSE
%    \STATE \textit{Deactivate reel motors}
%\ENDIF
%\end{algorithmic}
%\end{algorithm}

\subsection{Software Components}

\subsection{Control Strategy}

Two simple control strategies are proposed and evaluated.  The first one is called Follow-the-thread and the second strategy is Rotate-And-Go.  The Adafruit Shield controller provides a very stable output signal, provided battery are kept above reference.  Hence, wheels motor control is open-loop.

\section{Experimental Protocol} \label{experimental}

This section describes the experimental protocol used to evaluate the performance of the proposed solution.  The Pulmonary Rehabilitation procedure consists on a series of walking activities aimed to promote patient muscular recovery and well being~\cite{Wu2012}. They are slow pace motions following a specific trajectory on a rehabilitation gym.

In order to standardize the procedure~\cite{Sprunk2016}, the \textit{Lemniscate of Gerono} is used as desired trajectory, a curve shaped like an $\infty$ symbol, described by the Equation~\ref{eq:lemniscate}:

\begin{equation}
\begin{array}{ll}  
 x( \phi ) = a \; cos( \phi )     \\
 y( \phi ) = a \; cos ( \phi )   \; sin ( \phi )    \\ 
 \textit{where}  \; \phi \in \lbrace -\pi , \pi  \rbrace
\end{array}
\label{eq:lemniscate}
\end{equation}

\noindent where $ \phi $ is the free parameter, and $x$ and $y$ determine the shape of the trajectory on the navigation plane.

The reason  this shape was chosen is because it combines different kinds of trajectories that we want to test the vehicle in: long straight segments, sharp and soft curves, all in one single shape. Similar curves are also used in other proposed experiments in \cite{Neto2015,Endo2015}.  Regarding metrics, three are proposed to evaluate the performance.  They are

\begin{itemize}
    \item \textit{Normal path deviation}: the subject trajectory is divided into small segments and then the normal distance to the robot trajectory is calculated for each of those segments. Path deviation is relevant to evaluate how closely the robot mimics the leader path, which is the ultimate goal of the robotic vehicle. 
    \item \textit{Maximum path deviation}: the maximum \textit{normal path deviation} registered during an experiment.
    \item \textit{Total path deviation}: The area under the curve resulting from the the \textit{normal path deviation} over the length travelled by the leader.
    \item \textit{Robot-leader distance}: The euclidean distance between the robot and the leader, at any point in time. This is particularly important since the robot has a limited amount of thread available, so if the leader uses all the available thread, it will start dragging the robot and damaging the following mechanism. It is a scenario that must not happen under any circumstance, as it can also be dangerous for a potential patient using the device. 
\end{itemize}{}

In this work, a \textit{following behavior} is considered satisfactory if its maximum path deviation is less than 0.75 m and the robot-leader distance never exceeds 1.5 m~\cite{MunozCeballos2010}.

First the simulation is described and later the evaluation on the robotic prototype is detailed.

\subsection{Simulation}

A model of the proposed design was first built on Webots application.  The threading mechanism was implemented using virtual threads~\cite{Rekleitis2001}.  The leader traveled according to a predefined path with constant velocity, followiing the lemniscata trajectory.

The simulation is also useful to study the effects of the different constants in each strategy on the movement of the robot.

\subsection{Real world}

%To verify the validity of the proposed framework and method,
%The experimental protocol used to 
%In order to assess

A real world experiment was performed, begging to maintain the same conditions implemented on the simulation environment.  A lemniscata of $  a = 1 m $ was drawn on the floor on the visual capturing room.

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/capture.png}
\caption{Trajectory for the Follow-The-Thread control strategy.}
\label{fig:distance_sim}
\end{figure}

\section{Results} \label{Results}
\label{results}

Simulation results for both control strategies are shown on Figures 1 and 2.

Results metrics for the simualtions are shown on table 1.

\begin{center}
\begin{tabular}{ |c|c|c|c| }
\hline
$c_v$ & $c_{\alpha}$ & Max path deviation [m] & Area under \textit{npd} curve \\
\hline
10  &   15  & 0.3614 & 2.0651\\
15  &   5  & 0.4325 & 2.055\\
\textbf{15}  &   \textbf{10}  & \textbf{0.2188} & \textbf{1.0902}\\
15  &   15  & 0.2891 & 1.5059\\
5  &   20  & 0.5733 & 3.7289\\
\hline
\end{tabular}
\captionof{table}{Normal path deviation and maximum path deviation for different Follow the Thread constants.}
\label{tab:ft_npd_table}
\end{center}

\begin{center}
\begin{tabular}{ |c|c|c|c|c| }
\hline
$c_v$ & $c_r$ & $Dt_{off}$ [m] & Max path deviation [m] & Area under \textit{npd} curve \\
\hline
10  &   20  &   0.1  & 0.4310 & 1.6380\\
20  &   20  &   0.05  & 0.7775 & 3.1139\\
\textbf{20}  &  \textbf{20}  & \textbf{0.1} & \textbf{0.4123} & \textbf{0.9872}\\
20  &   35  &   0.1  & 0.4143 & 1.4820\\
20  &   5  &  0.05  & 0.7815 & 3.0892\\
35  &   20  &   0.1  & 0.6337 & 1.6190\\
\hline
\end{tabular}
\captionof{table}{Normal path deviation and maximum path deviation for different Rotate and Go constants.}
\label{tab:ft_npd_table}
\end{center}

Additionally, speed profiles can be observed from both trajectories.

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/ft_a2_n1500_cv15_cr15.eps}
\caption{Trajectory for the Follow-The-Thread control strategy.}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/rgv_cv20_cr20_dr10.eps}
\caption{Trajectory for the Rotate-And-Go control strategy.}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/ft_cv15_cr10_leader_robot_speed.eps}
\caption{Follow-The-Thread.}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/rg_cv20_cr20_bv10_dtoff10_leader_robot_speed.eps}
\caption{Speed profiles for Rotate-And-Go.}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/ft_vs_rg_dist.eps}
\caption{Distance between robot and leader [m]}
\label{fig:distance_sim}
\end{figure}

Results for the real world experiment can be seen on Figures 3 and 4.

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/ft1cap.eps}
\caption{Trajectories for the Follow-The-Thread control strategy.}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/rg2cap.eps}
\caption{Distance between robot and leader [m]}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/ft1cap_leader_robot_speed.eps}
\caption{Follow the thread speed profiles}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/rg2cap_leader_robot_speed.eps}
\caption{Rotate and Go speed profiles}
\label{fig:distance_sim}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=8cm]{images/cap_ft_vs_rg_dist.eps}
\caption{Rotate and Go speed profiles}
\label{fig:distance_sim}
\end{figure}

\begin{figure} 
    \centering
  \subfloat[a\label{1a}]{%
       \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
    \hfill
  \subfloat[b\label{1b}]{%
        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
    \\
  \subfloat[c\label{1c}]{%
        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
    \hfill
  \subfloat[d\label{1d}]{%
        \includegraphics[width=0.45\linewidth]{images/cap_ft_vs_rg_dist.eps}}
  \caption{(a), (b) Some examples from CIFAR-10 \cite{4}. The objects in     
        single-label images are usually roughly aligned.(c),(d) However, the 
        assumption of object alignment is not valid for multi-label
        images. Also note the partial visibility and occlusion
        between objects in the multi-label images.}
  \label{fig1} 
\end{figure}

\begin{center}
\begin{tabular}{ |c|c|c|c| }
\hline
$c_v$ & $c_{\alpha}$ & Max path deviation [m] & Area under \textit{npd} curve \\
\hline
25  &   20  & 0.3876 & 1.9761\\
25  &   35  & 0.4672 & 2.3528\\
\hline
\end{tabular}
\captionof{table}{Normal path deviation and maximum path deviation in motion capture experiments using Follow the Thread.}
\label{tab:cap_ft_npd_table}
\end{center}

\begin{center}
\begin{tabular}{ |c|c|c|c|c| }
\hline
$c_v$ & $c_r$ & $Dt_{off}$ & Max path deviation [m] & Area under \textit{npd} curve \\
\hline
30  &   35  & 0.04  & 0.4116 & 2.8309\\
30  &   35  & 0.08 & 0.3739 & 2.0367\\
\hline

\end{tabular}
\captionof{table}{Normal path deviation and maximum path deviation in motion capture experiments using Rotate and Go.}
\label{tab:cap_rg_npd_table}
\end{center}

\section{Discussion}
\label{discussion}

\subsection{Clinical Assessment} 
\label{clinical}

No amount of metrics are enough to evaluate if the robot is a viable solution for this problem or not, without the input and the evaluation of the people that are going to physically make use of it.

As this project is a joint research between ALPI and ITBA, we invited 4 people from the ALPI staff, 3 doctors and one engineer, to the motion capture laboratory at ITBA, where we made a live demonstration of the robot working and following a moving person.

In the demonstration, we described how the robot worked, how it was built and how to operate it. Both control strategies were explained, along with the main superficial differences between them.

Afterwards, we invited the doctors to use the robot themselves, simulating they were the patient being followed. We let the doctors switch between the two control strategies to evaluate both of them, and made different tests, one walking along path drawn in the floor (figure \ref{fig:cap_lab_lemniscate}), and another one walking freely in the space available in the lab. We let them use the robot as long as they wanted to, to get a general idea of how it behaved, and how it could be used in the rehabilitation process. After all evaluations were finished, we answered questions and discussed various aspects of the vehicle prototype with the ALPI staff, and then proceeded to ask them a survey to document their experience with the robot, get their expert opinion on how the two strategies compared against each other, and what other improvements were needed in order to deliver a fully usable product.

\begin{table*}[t]
\begin{center}
\begin{tabular}[!t]{|c|c|}
\hline
Question & Avg. answer \\
\hline
How would you qualify, from 1 to 5, your overall experience with Follow the Thread? (1:Bad, 5:Excellent)  & 5.0\\
\hline
How safe would a patient be, from 1 to 5, being followed by the robot using the Follow the Thread strategy? (1:Very unsafe, 5:Very safe)  & 4.25\\
\hline
How would you qualify, from 1 to 5, your overall experience with Rotate and Go? (1:Bad, 5:Excellent)  & 3.5\\
\hline
How safe would a patient be, from 1 to 5, being followed by the robot using the Rotate and Go strategy? (1:Very unsafe, 5:Very safe) & 3.25\\
\hline
\end{tabular}
\captionof{table}{Answers to survey questions.}
\label{tab:alpi_q_table_1}
\end{center}
\end{table*}

According to their answers, and the discussion we had after testing the robot, the general opinion was that the \textit{Follow the Thread} strategy was safer and more convenient for the task. In the survey, when asked \textit{Which of the two strategies is more effective at following the patient in a rehabilitation exercise?}, all 4 people responded that \textit{Follow the Thread} is "much better".

The main concern with the \textit{Rotate and Go} strategy was that having to wait for the robot to rotate before moving forward might be unsafe, as the patient could move away from it and compromise the cannula connecting him or her to the oxygen tank. This issue was identified during our own tests, and was not mentioned when explaining the following mechanism to the doctors, to avoid skewing them. They independently identified this problem, and emphasized that it could be a great source of discomfort for the patient.

Another aspect that was value from the \textit{Follow the Thread} strategy is that since its smoother movement, with no sudden stops or accelerations, was favorable for the stability of the robot if it were to carry the heavy oxygen tank.

Two needed security measures were also brought up by the ALPI team. Firstly, the need to add some mechanism for obstacle avoidance. They mentioned the need to have sensor to detect if the robot was about to hit something (specially the patient), and stop immediately, apart from what the control strategy indicated. Secondly, they recognized that some patients have very weak stability, and might fall down or take a step back, towards the robot, so it should be able to automatically move away from the patient, in order not to become another obstacle for him or her. "Tener en cuenta la distancia a la que se encuentra el robot por si el paciente llega a dar un paso atrás - para evitar riesto de caída'" and "Estaría bueno pensar en algún método de emergencia por caída'' were two quotes related to this issues, taken from the surveys.

In order to have more information for the next steps in the development of the robot, we asked for their advise to design the mechanism to attach the threads to the patient being followed. Two ideas were proposed: a belt strapped to the patient waist, or a clasp that would also to the clothes of the patient, also near its waistline. The waist is a good attachment point, since it is relatively stable when the patient moves, compared to its hands or legs, that may make sudden movements and confuse the robot sensors.

\subsection{Conclusion and Future Work}
\label{conclusion}

The goal of this work is to verify if a tethered robot to implement a following scheme on a patient during a pulmonary rehabilitation procedure is too simplistic to be a factual implementation.

From the practical experiments we showed that the both algorithms gave reasonable following performance in task of following the leader along a lemniscate-shaped path. This research is that this performance was achieved using very low cost components. It was proven that a reasonable following behaviour could be accomplished with a simple mechanism, a characteristic that significantly lowers the price of a device like this, putting it within reach of many medical institutions.

Each control strategy has its advantages, but according various metrics described in this work, the \textit{Follow and Turn} strategy had a more desirable behaviour, as it tended to follow the leader from a closer distance at all times, while moving in a smooth and predictable way.

We understood that no metric is enough to evaluate how appropriate the robotic vehicle can be for a real life scenario. Therefore, we gathered feedback from healthcare professionals from ALPI, who provided us with invaluable feedback to evaluate our solution. Over all, they also valued the Follow and Turn strategy as being the safer and more effective one. Most importantly, they also validated the research made and approved the direction of the project. They proposed a series of improvements and next steps, and after seeing the prototype in action, believed that we were close to a real, working solution.

It is essential to involve stakeholders such as patients, doctors, nurses, and any other professionals involved in the rehabilitation process early in the design roadmap. They are the ones who understand the problem better than anyone else, and will be the end users of a product like this.

The staff at ALPI was enthusiastic about helping in the development of the robot, and their encouragement and support were the key reasons we could take this project forward. This work not only made progress from a technical standpoint, but could also become the foundation of a long-term relationship of collaboration and research between ALPI and ITBA. This is also the start of a long-term project at ITBA. One that can go on to become a complete, real world solution that can improve the quality of life of many people through the use of technology and engineering.

\subsection{Future Work}
A following mechanism has been designed and implemented in a hardware prototype. The mechanism, along with the two control algorithms, has shown promise to be a viable solution to implement in a functioning robot that ALPI, or any other medical institution, can use in their every day exercises with COPD patients.

The next steps for this project would be to take this prototype and build a bigger version, with motors powerful enough to carry a real oxygen tank.

Further research is also needed to develop a full fledged solution: 
\begin{itemize}
    \item Power the motor wheels with a \textit{closed-loop and more efficient controller}, to have a more precise control over their speed.
    \item Redesign the active spring mechanism in the reels to avoid overheating. In our tests, the small electric motors in each reel got very hot after just 20 minutes of use. The algorithm we designed to use them more efficiently (\ref{alg:active_spring}) was enough to keep them in manageable temperatures (under $70^0C$), which was enough for testing the prototype, but would not be appropriate for a useful robot. Not only because the heat produced can be dangerous, but also because the motors can overheat and break their insulation in regular use.
    \item An easy and safe interaction between the patient, the operator and the robot. How to communicate the state of the robot to the operator, how to control and manipulate the robot in an effective and user-friendly way.
    \item Safety measures to keep the patient and the doctor or nurse safe when using the robot. Not only safe from the robot movement, but also from its electronic components, which can be dangerous when used near the pure oxygen tank if they are not protected properly.
    \item Investigate and design mechanisms for obstacle avoidance. This need was emphasized by the doctors from ALPI. The robot should have mechanisms in place to deal with emergency situations, and under no circumstance it can hit the patient or the doctor operating it.
    \item Achieve a battery autonomy that makes the robot useful throughout a complete pulmonary rehabilitation exercise. It is crucial for its usefulness to be able to hold a charge for this period of time, along with the ability to quickly swap batteries if the vehicle will be continually used with different patients.
\end{itemize}{}

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% use section* for acknowledgment
\section{Acknowledgments}
This project was supported by ALPI organization.

\section*{Conflict of Interest Statement}
The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
\bibliography{alpibot}

%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)\begin{IEEEbiographynophoto}
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{ramele}}]{Rodrigo Ramele}
%currently works at the Centro de Inteligencia Computacional (CIC), ITBA University, Buenos Aires, Argentina. Rodrigo does research in Brain Computer Interfaces, Artificial Intelligence, Signal Processing and Robotics.
%\end{IEEEbiography}
%
%% if you will not have a photo at all:
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{santos}}]{Juan Miguel Santos}
%is the Director of the Centro de Inteligencia Computacional (CiC) of ITBA University in Buenos Aires, Argentina, where he works in three different areas: robotics development, reinforcement learning and brain-computer interfaces.  In the first two lines he works since 1993 developing and constructing robots, national and international robotics conferences and international peer-review journals.  Currently developing a new development line based on SLAM to autonomous robots.
%\end{IEEEbiography}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}